Gozde and Liora Extension Project

### 2015-05-11 extension.Rmd

---
title: "Gozde and Liora extension"
author: "Liora and Gozde"
date: "April 27, 2015"
output:
  html_document:
    pandoc_args: [
      "+RTS", "-K64m",
      "-RTS"
    ]
---
# EXTENSION

@@

library(MASS)
library(nlme)
library(xtable)
library(foreign)
library(dplyr)
library(broom)
library(ggplot2)
library(xtable)
library(stargazer)
library(grid)
library(gridExtra)

@@

# Cleaning up data
data <- rename(data, ext_vote = rexvote)
data <- rename(data, age_1 = age1) # (18–29 years)
data <- rename(data, age_2 = age2) # (30–45 years)
data <- rename(data, age_4 = age4) # (>65 years)
data <- rename(data, edu_1 = mye1) # Education: middle/high
data <- rename(data, edu_2 = mye2) # Education: university
data <- rename(data, farmer_own = farmerown) 
data <- rename(data, left_right_scale = zlrs) # Left-Right Scale
data <- rename(data, eu_neg = euschlecht) # Negative evaluation of EU membership of one's own country
data <- rename(data, z_dem_satis = zsatisdmo) # Dissatisfied: Democracy
data <- rename(data, dispro_elect = disp) # Disproportionality
data <- rename(data, federalism = lfed1) # Decentralization
data <- rename(data, z_asylumseeker = zasylumseekers) 
data <- rename(data, z_unemp = zsur) # Unemployment rate
data <- rename(data, z_replacement = zreplacementrate) # Unemp benefits
data <- rename(data, max_er = rmax) # Maximum toughness in extreme right discourse

@@

# Model 1: Adorno Thesis (Personality Traits and Value Orientations)
The first model is based on Theodor Adorno's 1950 theory that support for extreme right parties (in his case, fascist parties) originates from supporters' personality traits and value orientations, which make individuals more receptive to the appeals of extreme right parties.  We operationalize "personality traits and value orientations" using the individual level characteristics from our data set: gender, age, occupation, marital status, and self-placement on a left-right scale.

```{r}
library(separationplot)
library(Hmisc)

model_1 <- glm(ext_vote ~ male + age_1 + age_2 + age_4 + edu_1 + edu_2 + farmer_own + worker + retired + unemployed + partner + left_right_scale + country - 1, data = data)
summary(model_1)
actual <- model_1$y # Create vector with actual values for vote intention for extreme right parties

log.odds <- model_1$fitted.values # Create a vector with log-odd values
# Create link function to transform log-odds into predicted probabilities
link <- function(P) {
  1/(1+exp(-P))
} 
pred <- link(log.odds) # Transform log-odds into predicted probabilities
pred <- as.vector(pred) 
range(pred) # variation in the predicted probabilities: very little variation.
dev.off()
# Create separation plot to visualize the predictive power of the model
separationplot(pred = pred, actual = actual, type = "rect", heading = "Model1: The Adorno Thesis", xlab = "ER Vote Intention", newplot = F, line = T, lwd2 = 0.8, height = 0.5, show.expected = T) 
```

@@

actual <- model_2$y # Create vector with actual values for vote intention for extreme right parties

log.odds <- model_2$fitted.values # Create a function with log-odds
# Create link function to transform log-odds into predicted probabilities
link <- function(P) {
  1/(1+exp(-P))
} 
pred <- link(log.odds) # Transform log-odds into predicted probabilities
pred <- as.vector(pred) 
range(pred) # variation in the predicted probabilities: very little variation.
dev.off()
# Create separation plot to visualize the predictive power of the model
separationplot(pred = pred, actual = actual, type = "rect", heading = "Model2: The Parsons Thesis", xlab = "ER Vote Intention", newplot = F, line = T, lwd2 = 0.8, height = 0.5, show.expected = T) 
```

@@

#Conclusion:
Again, there is little variation in our predictive probabilities with this model. Similar to model 1, most events are aligned with higher predictive probabilities on the right hand of the graph. Yet Model 1 has slightly better predictive power as the very end of the graph is likely to be clustered with event than non-events as in Model 2.

@@

@@

log.odds <- model_3$fitted.values # Create a function with log-odds
# Create link function to transform log-odds into predicted probabilities
link <- function(P) {
  1/(1+exp(-P))
} 
pred <- link(log.odds) # Transform log-odds into predicted probabilities
pred <- as.vector(pred) 
range(pred) # variation in the predicted probabilities: very little variation.
dev.off()
# Create separation plot to visualize the predictive power of the model
separationplot(pred = pred, actual = actual, type = "rect", heading = "Model3: The Group Conflict Thesis", xlab = "ER Vote Intention", newplot = F, line = T, lwd2 = 0.8, height = 0.5, show.expected = T) 
```

@@

#Conclusion:
The majority of events (intention to vote for ER) are clustered at the end right end of the model - which is in line with a good predictive model. However, in this model the distribution of events is very dispersed compared to other two models.

>>>>

---
title: "Gozde and Liora Extension Project"
author: "Liora and Gozde"
date: "April 27, 2015"
output:
  html_document:
    pandoc_args: [
      "+RTS", "-K64m",
      "-RTS"
    ]
---

@@

library(MASS)
library(nlme)
library(foreign)
library(dplyr)
library(broom)
library(ggplot2)
library(stargazer)
library(grid)
library(gridExtra)
library(separationplot)
library(Hmisc)

@@

# Cleaning up data
data <- data %>%
  rename(ext_vote = rexvote,
         age_1 = age1, # (18–29 years)
         age_2 = age2, # (30–45 years)
         age_4 = age4, # (>65 years)
         edu_1 = mye1, # Education: middle/high
         edu_2 = mye2, # Education: university
         farmer_own = farmerown,
         left_right_scale = zlrs, # Left-Right Scale
         eu_neg = euschlecht, # Negative evaluation of EU membership of one's own country
         z_dem_satis = zsatisdmo, # Dissatisfied: Democracy
         dispro_elect = disp, # Disproportionality
         federalism = lfed1, # Decentralization
         z_asylumseeker = zasylumseekers,
         z_unemp = zsur, # Unemployment rate
         z_replacement = zreplacementrate, # Unemp benefits
         max_er = rmax) # Maximum toughness in extreme right discourse
```

@@

# Model 1: Adorno Thesis (Individual Characteristics and Value Orientations)
The first model is based on Theodor Adorno's 1950 theory that support for extreme right parties (in his case, fascist parties) originates from supporters' personality traits and value orientations, which make individuals more receptive to the appeals of extreme right parties.  We operationalize "personality traits and value orientations" using the individual level characteristics from our data set: gender, age, occupation, marital status, and self-placement on a left-right scale.

```{r}
model_1 <- glm(ext_vote ~ male + age_1 + age_2 + age_4 + edu_1 + edu_2 + farmer_own + worker + retired + unemployed + partner + left_right_scale + country - 1, data = data)
summary(model_1)
actual <- model_1$y # Create vector with actual values for vote intention for extreme right parties

log.odds <- model_1$fitted.values # Create a vector with log-odd values
# Create link function to transform log-odds into predicted probabilities
link <- function(log.odd) {
  pred.prob <- 1/(1+exp(-log.odd))
  pred.prob # returns a numeric vector of predicted probabilities
} 
pred_prob <- link(log.odds) # Transform log-odds into predicted probabilities
pred_prob <- as.vector(pred_prob) 
range(pred_prob) # variation in the predicted probabilities: very little variation.
# Graph variation in Predicted probabilities
pred_df <- data.frame(pred_prob)
ggplot(pred_df, aes(x = 1:length(pred_prob), y = pred_prob)) + 
  geom_line() +
  xlab("Prediction Number") +
  ylab("Predicted Probability") +
  ggtitle("Model 1: Variation in Predicted Probabilities")
# Create separation plot to visualize the predictive power of the model
separationplot(pred = pred_prob, actual = actual, type = "rect", heading = "Model1: The Adorno Thesis", xlab = "ER Vote Intention", newplot = F, line = T, lwd2 = 0.8, height = 0.5, show.expected = T) 
```

@@

actual <- model_2$y # Create vector with actual values for vote intention for extreme right parties
log.odds <- model_2$fitted.values # Create a function with log-odds
pred_prob <- link(log.odds) # Transform log-odds into predicted probabilities
pred_prob <- as.vector(pred_prob) 
range(pred_prob) # variation in the predicted probabilities: very little variation.
# Visualize variation in predicted probabilities
pred_df <- data.frame(pred_prob)
ggplot(pred_df, aes(x = 1:length(pred_prob), y = pred_prob)) + 
  geom_line() +
  xlab("Prediction Number") +
  ylab("Predicted Probability") +
  ggtitle("Model 2: Variation in Predicted Probabilities")
# Create separation plot to visualize the predictive power of the model
separationplot(pred = pred_prob, actual = actual, type = "rect", heading = "Model2: The Parsons Thesis", xlab = "ER Vote Intention", newplot = F, line = T, lwd2 = 0.8, height = 0.5, show.expected = T) 
```

@@

#Conclusion:
Again, there is little variation in our predictive probabilities with this model. Similar to model 1, most events are aligned with higher predictive probabilities on the right hand of the graph. Yet Model 1 has slightly better predictive power as the very end of the graph is likely to be clustered with events rather than non-events as in Model 2.

@@

We used interactions between contextual variables in this model in line with the author's original model.

@@

log.odds <- model_3$fitted.values # Create a function with log-odds
pred_prob <- link(log.odds) # Transform log-odds into predicted probabilities
pred_prob <- as.vector(pred_prob) 
range(pred_prob) # variation in the predicted probabilities: very little variation.
# Visualize variation in predicted probabilities
pred_df <- data.frame(pred_prob)
ggplot(pred_df, aes(x = 1:length(pred_prob), y = pred_prob)) + 
  geom_line() +
  xlab("Prediction No") +
  ylab("Predicted Probability") +
  ggtitle("Model 3: Variation in Predicted Probabilities")
# Create separation plot to visualize the predictive power of the model
separationplot(pred = pred_prob, actual = actual, type = "rect", heading = "Model3: The Group Conflict Thesis", xlab = "ER Vote Intention", newplot = F, line = T, lwd2 = 0.8, height = 0.5, show.expected = T) 
```

@@

#Conclusion:
The majority of events (the intention to vote for an extreme right party) is clustered at the right end of the model - which is in line with a good predictive model. However, in this model the distribution of events is very dispersed compared to the other two models.
